Toward single particle reconstruction without particle picking: Breaking the detection limit

### 2018-09-02 pnas/Towards.tex
Files sent to Amit

\begin{figure*}[h!]
	\centering
	\includegraphics[scale=0.7]{1D_example}
	\caption{\label{fig:1D_example} Recovery of a 1-D signal of length $L=21$ at noise level $\sigma = 3$ from $M$ repetitions. The length of the micrograph was set to be $N = 410*M$. The first three panels (left to right) show reconstruction with different $M$ values compared to the ground truth signal (in blue).
		The true value of $\gamma$ is $0.0512$.
		   The relative errors of $\gamma$ for the three panels are: $27.1\%,4\%,1.2\%$. The right panel shows the relative error as a function of $M$.  \TODO{Nico: relative error should be logscale? Also, let's pick other colors; if we're running this again anyway, let's also aim for M being powers of 10 in all displayed cases? To run with logspace(3,7,9). We can remove the x-axis}
	}
\end{figure*}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}[h]{0.45\textwidth}
		\centering
		\includegraphics[angle=270, width=.7\textwidth]{TRPV1_rec.eps}
		\caption{TPRV1}\label{fig:TRPV1_recon}
	\end{subfigure}
	\hfill
	\begin{subfigure}[h]{0.45\textwidth}
		\centering
		\includegraphics[angle=270, width=.7\textwidth]{BPTI_maxL2_L20.eps}
		\caption{BPTI}\label{fig:BPTI_recon}
	\end{subfigure} 
	\caption{\label{fig:cryo_recon} Reconstructions from the first 3 autocorrelations. The reconstruction is shown in gray while the ground truth is shown in yellow.}
\end{figure}

@@

The true volume used in the experiments in Figures~\ref{fig:cryo_detection} and \ref{fig:BPTI_recon} was the Bovine Pancreatic Trypsin Inhibitor (BPTI) mutant with altered binding loop sequence, whos atomic model is available in the Protein Data Bank (PDB) %~\cite{pdb} 
as 1QLQ\footnote{\url{https://www.rcsb.org}}. We generated an EM map from this atomic model in UCSF Chimera~\cite{pettersen2004ucsf} at a resolution of 5 \AA, and cropped it to remove zeros at the boundary to obtain a volume of size $31^3$. For the experiment in Figure~\ref{fig:BPTI_recon}, the volume was downsampled to size $20^3$. For the experiment in Figure~\ref{fig:TRPV1_recon}, we used the TRPV1 in complex with DkTx and RTX, whose EM map is available in the  Electron Microscopy Data Bank (EMDB) %~\cite{emdb} 
as EMD-8117\footnote{\url{http://www.ebi.ac.uk/pdbe/emdb}}. The original map has size $192^3$, and was downsampled to size $20^3$. To generate the ground truth for our reconstructions, we expanded both volumes as in \eqref{eq:volume_expansion_app} with cutoff $L = 5$ for TRPV1 and $L = 2$ for BPTI \TODO{Do we need to explain why?}. For the TRPV1 reconstruction, the optimizer converged to a point giving relative $\ell_2$ error of $10^{-6}$ in the first three autocorrelations and an error of $10^{-1}$ in the expansion coefficients of the volume. For the BPTI reconstruction, the errors in the autocorrelations were $10^{-6}$, $10^{-7}$ and $10^{-8}$ for the third, second and first autocorrelations, respectively, while the error in the expansion coefficients of the volume was $5\times 10^{-2}$. This illustrates the ill-conditioning of the map between the volume and its first three autocorrelations that prevents us from obtaining results from noisy data. \TODO{Tamir: This entire paragraph will be changed a bit once we will have the final figures}

@@

\section{Proof of Proposition~\ref{prop:gamma_sigma}} \label{sec:proof_prop_gamma_sigma}

We will prove that both $\sigma$ and $\gamma$ are identifiable from the observed first three moments of $y$. For convenience, we will work with $\beta = \gamma / L$ rather than $\gamma$ itself. We will construct two quadratic equations satisfied by $\beta$ from observed quantities, independent of $\sigma$. Then, we will show that these equations are independent, and hence that $\beta$ is uniquely defined.  Given $\beta$, we can estimate $\sigma$ using Proposition~\ref{prop:gamma}.

>>>>

\begin{figure*}[h!]
	\centering
	\includegraphics[scale=0.7]{1D_example}
	\caption{\label{fig:1D_example} Recovery of a 1-D signal of length $L=21$ at noise level $\sigma = 3$ from $M$ repetitions. The length of the micrograph was set to be $N = 410*M$. The first three panels (left to right) show reconstruction with different $M$ values compared to the ground truth signal (in blue).
		The true value of $\gamma$ is $0.0512$.
		   The relative errors of $\gamma$ for the three panels are: $4.8\%,4\%,1.2\%$. The right panel shows the relative error as a function of $M$. 
	}
\end{figure*}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}[h]{0.45\textwidth}
		\centering
		\includegraphics[angle=270, width=.7\textwidth]{TRPV1_rec.eps}
		\caption{TPRV1 with cutoff $L=5$}\label{fig:TRPV1_recon}
	\end{subfigure}
	\hfill
	\begin{subfigure}[h]{0.45\textwidth}
		\centering
		\includegraphics[angle=270, width=.7\textwidth]{BPTI_maxL2_L20.eps}
		\caption{BPTI with cutoff $L=2$}\label{fig:BPTI_recon}
	\end{subfigure} 
	\caption{\label{fig:cryo_recon} Reconstructions from the first 3 autocorrelations.
		The ground truth volumes were expanded according to~\eqref{eq:volume_expansion_app} with cutoff $L$.
		The reconstruction is shown in gray while the (smoothed) ground truth in yellow. }
\end{figure}

@@

The true volume used in the experiments in Figures~\ref{fig:cryo_detection} and~\ref{fig:BPTI_recon} was the Bovine Pancreatic Trypsin Inhibitor (BPTI) mutant with altered binding loop sequence, whos atomic model is available in the Protein Data Bank (PDB) %~\cite{pdb} 
as 1QLQ\footnote{\url{https://www.rcsb.org}}. We generated an EM map from this atomic model in UCSF Chimera~\cite{pettersen2004ucsf} at a resolution of 5~\AA, and cropped it to remove zeros at the boundary to obtain a volume of size $31^3$. For the experiment in Figure~\ref{fig:BPTI_recon}, the volume was downsampled to size $20^3$. For the experiment in Figure~\ref{fig:TRPV1_recon}, we used the TRPV1 in complex with DkTx and RTX, whose EM map is available in the  Electron Microscopy Data Bank (EMDB) %~\cite{emdb} 
as EMD-8117\footnote{\url{http://www.ebi.ac.uk/pdbe/emdb}}. The original map has size $192^3$, and was downsampled to size $20^3$. To generate the ground truth for our reconstructions, we expanded both volumes as in~\eqref{eq:volume_expansion_app} with cutoff $L = 5$ for TRPV1 and $L = 2$ for BPTI. For the TRPV1 reconstruction, the optimizer converged to a point giving relative $\ell_2$ error of $10^{-6}$ in the first three autocorrelations and an error of $10^{-1}$ in the expansion coefficients of the volume. For the BPTI reconstruction, the errors in the autocorrelations were $10^{-6}$, $10^{-7}$ and $10^{-8}$ for the third, second and first autocorrelations, respectively, while the error in the expansion coefficients of the volume was $5\times 10^{-2}$. This illustrates the ill-conditioning of the map between the volume and its first three autocorrelations that prevents us from obtaining results from noisy data.

@@

\section{Proof of Proposition~\ref{prop:gamma_sigma}} \label{sec:proof_prop_gamma_sigma}
\TODO{Need to check the details of the proof one more time}
We will prove that both $\sigma$ and $\gamma$ are identifiable from the observed first three moments of $y$. For convenience, we will work with $\beta = \gamma / L$ rather than $\gamma$ itself. We will construct two quadratic equations satisfied by $\beta$ from observed quantities, independent of $\sigma$. Then, we will show that these equations are independent, and hence that $\beta$ is uniquely defined.  Given $\beta$, we can estimate $\sigma$ using Proposition~\ref{prop:gamma}.
